Reproducible Research Assignment 1
======================================================


```{r}
# Reading Data in
library(data.table)
library(knitr)

setwd("/Users/Robert/Desktop/DataScienceTrack/ReproducibleResearch")
#url <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

#download.file(url,paste(getwd(),"/assign1.zip",sep=""),"curl")
#unzip("assign1.zip")
data <- data.table(fread("activity.csv"))
```

```{r}
dayTotals <- tapply(data$steps,data$date,sum,na.rm=TRUE)
```
histogram of the total number of steps taken each day
```{r}
hist(dayTotals,main="Histogram of the total number of steps taken each day")
```
mean and median number of steps taken each day 
```{r}
mean(dayTotals,na.rm=TRUE)
median(dayTotals,na.rm=TRUE)
```

```{r}
minuteTotals <- tapply(data$steps,data$interval,sum,na.rm=TRUE)
plot(unique(data$interval),minuteTotals,type="l",ylab="Average number of steps taken",xlab="5 minute interval")
abline(h=mean(dayTotals,na.rm=TRUE))
```

5-minute interval that, on average, contains the maximum number of steps
```{r}
data$interval[which(minuteTotals==max(minuteTotals))]
```

Strategy for imputing missing data
```{r} 
nrow(data)-length(which(complete.cases(data)))

naI <- which(complete.cases(data)==FALSE)
data2 <- data
data2$steps[naI] <- minuteTotals[which(unique(data$interval)==data2$interval[naI])]
dayTotals2 <- tapply(data2$steps,data2$date,sum,na.rm=TRUE)
```

Histogram of the total number of steps taken each day after missing values were imputed
```{r}
hist(dayTotals2,main="Histogram of the total number of steps taken each day")
```


```{r}
day <- weekdays(as.Date(data$date))

weekendI <- which(day=="Sunday"|day=="Saturday")
day[1:17568] <- "weekday"
day[weekendI] <- "weekend"
day <- as.factor(day)
data3 <- cbind(data,day)

```
Panel plot comparing the average number of steps taken per 5- minute interval across weekdays and weekends
```{r}
minuteTotals2 <- tapply(data3$steps,list(data3$interval,data3$day),sum,na.rm=TRUE)
par("mfrow"=c(2,1))
plot(unique(data3$interval),minuteTotals2[,1],ylab="number of steps",xlab="interval",main="Weekdays",type="l")
plot(unique(data3$interval),minuteTotals2[,2],ylab="number of steps",xlab="interval",main="Weekend",type="l")
```